---
title: "EDA Project 1"
author: "Michael Burgess & Robert Pollock"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=6) 
```

```{r}
library(dplyr)
library(ggplot2)
library(cowplot)
library(class)
library(caret)
library(e1071)
```
##Introduction

We were asked by our clients, the CEO and CFO of Budweiser, to examine data related to different craft beers and breweries located throughout the United states. The data contains 2,410 beers and 558 breweries. In this presentation we will answer 7 questions that were asked by our clients.

```{r}
#data import
beers <- read.csv(choose.files(),header = T,sep = ",")
breweries <- read.csv(choose.files(),header = T,sep = ",")
```

##Question 1
How many breweries are present in each state?

The table below provides a brief look at the number of breweries per state and is arranged in alphabetical order.  

```{r}
table(breweries$State)
```

##Question 2
Merge beer data with breweries

```{r}
merged_df <- left_join(beers, breweries, by = c("Brewery_id" = "Brew_ID"))
names(merged_df)[names(merged_df) == "Name.x"] <- "Beer Name"
names(merged_df)[names(merged_df) == "Name.y"] <- "Brewery Name"
final_df <- merged_df[order(merged_df[,10], merged_df[9], merged_df[8], merged_df[1]),]

head(final_df,6)
tail(final_df,6)
```

##Question 3
Address the missing values in each column.

Once the 2 data sets were joined, we found that there were 62 beers with missing values for ABV and 1005 beers missing data for IBU. There were no missing values in the other columns. To fill in the missing ABV and IBU values we took the average values of ABV and IBU for each style of beers and used these mean values to fill in the missing data. There are 52 beers that continued to have missing IBU values after filling in the mean values. We found that these are either styles that only occur once or twice out of over 2000 different beers or are beers we would expect to be very sweet such as ciders or meads. We felt comfortable that leaving out these beers in our analysis would not greatly skew the data or our results. The data should be used with caution as outliers could greatly affect the mean values and we know that craft breweries do like to experiment with different ABV and IBU values, even within one style of beer. 

```{r}
#find mean ABV and mean IBU for each style and store it in a dataframe
mean_ABV <- final_df %>% filter(!is.na(ABV)) %>% group_by(Style) %>% summarize(meanABV = mean(ABV), count = n()) %>% arrange(meanABV)
mean_IBU <- final_df %>% filter(!is.na(IBU)) %>% group_by(Style) %>% summarize(meanIBU = mean(IBU), count = n()) %>% arrange(meanIBU)

#join the mean ABV rating to the main dataframe and drop count
final_df <- left_join(final_df, mean_ABV, by = "Style")
final_df <- final_df[,-c(12)]

#join the mean IBU rating to the main dataframe and drop count
final_df <- left_join(final_df, mean_IBU, by = "Style")
final_df <- final_df[,-c(13)]

#find row number of na values and store them
abv_na_rows <- which(is.na(final_df$ABV) == T)

#replace NA ABV values  with mean ABV values
for(i in abv_na_rows){
  final_df$ABV[i] <- final_df$meanABV[i]
}

#find row numbers of na values for ibu and store them
ibu_na_rows <- which(is.na(final_df$IBU) == T)

#replace NA IBU values with mean IBU values
for(i in ibu_na_rows){
  final_df$IBU[i] <- final_df$meanIBU[i]
}

#check to ensure that missing IBU values were not caused by code error
missing_IBU <- final_df %>% filter(is.na(final_df$IBU))
table(missing_IBU$Style)

#remove meanABV and meanIBU columns since they are no longer needed
final_df <- final_df[,-c(11, 12)]
```

##Question 4
Plot the median alcohol content and IBU for each state. Plot a bar chart to compare.

We found the median (middle) value for ABV and IBU for each state. The plots below show how each state compares to the others for both their median ABV and IBU. 

```{r}
median_ABV <- final_df %>% group_by(State) %>% summarize(medianABV = median(ABV), count = n()) %>% arrange(State)
median_IBU <- final_df %>% filter(!is.na(final_df$IBU)) %>% group_by(State) %>% summarize(medianIBU = median(IBU), count = n()) %>% arrange(State)

median_ABVPlot <- median_ABV %>% ggplot(mapping = aes(x=State, y=medianABV, fill = State)) + geom_col(show.legend = FALSE) + ylab("Median ABV") + ggtitle("Median ABV by State")
median_IBUPlot <- median_IBU %>% ggplot(mapping = aes(x=State, y=medianIBU, fill = State)) + geom_col(show.legend = FALSE) + ylab("Median IBU") + ggtitle("Median IBU by State")
plot_grid(median_ABVPlot, median_IBUPlot, labels = "AUTO", ncol = 1)
```

##Question 5
Which state has the maximum alcholic (ABV) beer? Which state has the most bitter (IBU) beer?

We found that the Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale from Upslope Brewing Company in Boulder, Colorado had the highest ABV at 12.8%. 
We found that the Bitter Birch Imperial IPA from Astoria Brewing COmpany in Astoria, Oregon had the highest IBU at 138. 

```{r}
final_df[which.max(final_df$ABV),]
final_df[which.max(final_df$IBU),]
```
##Question 6
Comment on the summary statistics and distribution of the ABV variable.

The summary shows that the minimum ABV is 0.1%, the maximum is 12.8%, and the median is at 5.65% which is about what we would expect to see. The mean is at 5.98%, so nearly the same as the median. This shows that there are not very many extreme outliers pulling the data in either direction. The histogram and bar plots show similar results with almost 450 beers at 5% ABV, a little less than 400 beers at around 6% ABV, and then a gradual tapering off towards higher ABV values. 
```{r}
summary(final_df$ABV)

abv_hist <- final_df %>% ggplot(mapping = aes(x=ABV)) + geom_histogram() + labs(y = "Count of Beers", title = "Distribution of ABV")
abv_box <- final_df %>% ggplot(mapping=aes(y=ABV)) + geom_boxplot() + labs(title = "Distribution of ABV") + theme(axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank()) 
plot_grid(abv_hist, abv_box, labels = "AUTO")

```
##Question 7
Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. Make your best judgment of a relationship and EXPLAIN your answer.

It appears that there is a weak relationship between IBU and ABV for beers with IBU values between 0 and 50. Most of the beers here are centered around 5% ABV with a  slight increase in ABV as IBU increases between 0 and 25 and then staying at a fairly consistent ABV for beers between 25 and 50 IBU. For beers with an IBU between 50 and 100 there appears to be more of a linear relationship with ABV increasing more consistently as IBU increases. This relationship tends to flatten out above 100 IBU, but there are fewer beers at this IBU level and this lack of a relationship may simply be a result of a lack of a data. 

```{r}
final_df %>% ggplot(mapping = aes(x = IBU, y = ABV)) + geom_point(position = "jitter") + geom_smooth(method ="loess") +ggtitle("Relationship Between IBU and ABV")
```
##Question 8
Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with "Ale" in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages . KNN is very easy to understand conceptually.

We ran tests using several different values of K, and found that a K value of 5 provided the best accuracy. Using IBU and ABV our model could accurately predict whether a beer was an IPA or Ale with approx 90% accuracy. It predicted ales correctly approx 92% of the time and IPAs approx 87% correctly. 

```{r}

final_df['Group'] <- NA

#find rows with IPA
ipa_rows <- grep("IPA", final_df$Style, ignore.case = T)

#replace NA in group with IPA for IPA rows
for(i in ipa_rows){
  final_df$Group[i] <- "IPA"
}

#find rows with Ale
ale_rows <- grep("Ale", final_df$Style, ignore.case = T)

#replace NA in group with Ale for Ale rows (the is.NA is necessary to avoid over writing IPA styles that also had ale in the name)
for(i in ale_rows){
  if(is.na(final_df$Group[i])==T){
    final_df$Group[i] <- "Ale"
  }
}

ABV_data <- final_df %>% filter(Group == "Ale" | Group == "IPA")
ABV_data <- ABV_data %>% filter(!is.na(IBU))
  
#run several tests to find best value of k
splitPerc = .70

iterations = 500
numks = 15

masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  accs = data.frame(accuracy = numeric(15), k = numeric(15))
  trainIndices = sample(1:dim(ABV_data)[1],round(splitPerc * dim(ABV_data)[1]))
  train = ABV_data[trainIndices,]
  test = ABV_data[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(3,4)],test[,c(3,4)],train$Group, prob = TRUE, k = i)
    CM = confusionMatrix(table(classifications,test$Group))
    masterAcc[j,i] = CM$overall[1]
  }
  
}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")

classifications = knn(train[,c(3,4)],test[,c(3,4)],train$Group, prob = TRUE, k = 5)
CM = confusionMatrix(table(classifications,test$Group))

CM
```

