---
title: "EDA Project 1"
author: "Michael Burgess & Robert Pollock"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=6) 
```

```{r}
library(dplyr)
library(ggplot2)
library(cowplot)
```
##Introduction

We were asked by our clients, the CEO and CFO of Budweiser, to examine data related to different craft beers and breweries located throughout the United states. The data contains 2,410 beers and 558 breweries. In this presentation we will answer 7 questions that were asked by our clients.

```{r}
#data import
beers <- read.csv(choose.files(),header = T,sep = ",")
breweries <- read.csv(choose.files(),header = T,sep = ",")
```

##Question 1

How many breweries are present in each state?

```{r}
table(breweries$State)
```

##Question 2
Merge beer data with breweries

```{r}
merged_df <- left_join(beers, breweries, by = c("Brewery_id" = "Brew_ID"))
names(merged_df)[names(merged_df) == "Name.x"] <- "Beer Name"
names(merged_df)[names(merged_df) == "Name.y"] <- "Brewery Name"
final_df <- merged_df[order(merged_df[,10], merged_df[9], merged_df[8], merged_df[1]),]

head(final_df,6)
tail(final_df,6)
```

#abv has 62 missing values and IBU has 1005 missing values for the merged DF, no other missing data or extra rows so join is fine
#for the NA values how to compute?
```{r}
#find mean ABV and mean IBU for each style and store it in a dataframe
mean_ABV <- final_df %>% filter(!is.na(ABV)) %>% group_by(Style) %>% summarize(meanABV = mean(ABV), count = n()) %>% arrange(meanABV)
mean_IBU <- final_df %>% filter(!is.na(IBU)) %>% group_by(Style) %>% summarize(meanIBU = mean(IBU), count = n()) %>% arrange(meanIBU)

#join the mean ABV rating to the main dataframe and drop count
final_df <- left_join(final_df, mean_ABV, by = "Style")
final_df <- final_df[,-c(12)]

#join the mean IBU rating to the main dataframe and drop count
final_df <- left_join(final_df, mean_IBU, by = "Style")
final_df <- final_df[,-c(13)]

#find row number of na values and store them
abv_na_rows <- which(is.na(final_df$ABV) == T)

#replace NA ABV values  with mean ABV values
for(i in abv_na_rows){
  final_df$ABV[i] <- final_df$meanABV[i]
}

#find row numbers of na values for ibu and store them
ibu_na_rows <- which(is.na(final_df$IBU) == T)

#replace NA IBU values with mean IBU values
for(i in ibu_na_rows){
  final_df$IBU[i] <- final_df$meanIBU[i]
}

#check to ensure that missing IBU values were not caused by code error
missing_IBU <- final_df %>% filter(is.na(final_df$IBU))
table(missing_IBU$Style)

#compare count of all styles to styles missing IBU: they match so the beers in missing_IBU had no values for all entries
table(final_df$Style)

#remove meanABV and meanIBU columns since they are no longer needed
final_df <- final_df[,-c(11, 12)]
```

##Plot the median alcohol content and IBU for each state. Plot a bar chart to compare.

```{r}
median_ABV <- final_df %>% group_by(State) %>% summarize(medianABV = median(ABV) * 100, count = n()) %>% arrange(State)
median_IBU <- final_df %>% filter(!is.na(final_df$IBU)) %>% group_by(State) %>% summarize(medianIBU = median(IBU), count = n()) %>% arrange(State)

median_ABVPlot <- median_ABV %>% ggplot(mapping = aes(x=State, y=medianABV, fill = State)) + geom_col(show.legend = FALSE) + ylab("Median ABV") + ggtitle("Median ABV by State")
median_IBUPlot <- median_IBU %>% ggplot(mapping = aes(x=State, y=medianIBU, fill = State)) + geom_col(show.legend = FALSE) + ylab("Median IBU") + ggtitle("Median IBU by State")
plot_grid(median_ABVPlot, median_IBUPlot, labels = "AUTO", ncol = 1)
```