---
title: "EDA Project 1"
author: "Michael Burgess & Robert Pollock"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=20, fig.height=6) 
```

```{r}
library(dplyr)
library(ggplot2)
library(cowplot)
library(class)
library(caret)
library(e1071)
```
##Introduction

We were asked by our clients, the CEO and CFO of Budweiser, to examine data related to different craft beers and breweries located throughout the United states. The data contains 2,410 beers and 558 breweries. In this presentation we will answer 7 questions that were asked by our clients.

```{r}
#data import
beers <- read.csv(choose.files(),header = T,sep = ",")
breweries <- read.csv(choose.files(),header = T,sep = ",")
```

##Question 1
How many breweries are present in each state?

The table and chart below provides a brief look at the number of breweries per state and is arranged in alphabetical order.  

```{r}
#create a dataframe to hold the frequency count of breweries in each state
state_brew_count <- as.data.frame(table(breweries$State))
#plot the count
state_count_plot <-state_brew_count %>% ggplot(mapping = aes(x = Var1, y = Freq)) + geom_col(fill = "#13294b") + 
  geom_text(aes(label = Freq), position=position_dodge(width=0.9), vjust=-0.25) + 
  labs(title = "Count of Craft Breweries By State", x = "State", y = "Number of Breweries")
#make the table more human readable
names(state_brew_count)[names(state_brew_count) == "Var1"] <- "State"
names(state_brew_count)[names(state_brew_count) == "Freq"] <- "Number of Breweries"
state_brew_count
state_count_plot
```

##Question 2
Merge beer data with breweries

```{r}
#I used left join in case any of the beers didn't have brewery info, but this wasn't an issue. The only columns with NA's were in the IBU and ABV and they were in the data from the beginning
merged_df <- left_join(beers, breweries, by = c("Brewery_id" = "Brew_ID"))
#make column names more readable
names(merged_df)[names(merged_df) == "Name.x"] <- "Beer Name"
names(merged_df)[names(merged_df) == "Name.y"] <- "Brewery Name"
#organize data by state, city, brewery name, and then beer name
final_df <- merged_df[order(merged_df[,10], merged_df[9], merged_df[8], merged_df[1]),]
head(final_df,6)
tail(final_df,6)
```

##Question 3
Address the missing values in each column.

Once the 2 data sets were joined, we found that there were 62 beers with missing values for ABV and 1005 beers missing data for IBU. There were no missing values in the other columns. To fill in the missing ABV and IBU values we took the average values of ABV and IBU for each style of beers and used these mean values to fill in the missing data. There are 52 beers that continued to have missing IBU values after filling in the mean values. We found that these are either styles that only occur once or twice out of over 2000 different beers or are beers we would expect to be very sweet such as ciders or meads. We felt comfortable that leaving out these beers in our analysis would not greatly skew the data or our results. The data should be used with caution as outliers could greatly affect the mean values and we know that craft breweries do like to experiment with different ABV and IBU values, even within one style of beer. 

```{r}
#find mean ABV and mean IBU for each style and store it in a dataframe
mean_ABV <- final_df %>% filter(!is.na(ABV)) %>% group_by(Style) %>% summarize(meanABV = mean(ABV), count = n()) %>% arrange(meanABV)
mean_IBU <- final_df %>% filter(!is.na(IBU)) %>% group_by(Style) %>% summarize(meanIBU = mean(IBU), count = n()) %>% arrange(meanIBU)
#join the mean ABV rating to the main dataframe and drop count
final_df <- left_join(final_df, mean_ABV, by = "Style")
final_df <- final_df[,-c(12)]
#join the mean IBU rating to the main dataframe and drop count
final_df <- left_join(final_df, mean_IBU, by = "Style")
final_df <- final_df[,-c(13)]
#find row number of na values and store them
abv_na_rows <- which(is.na(final_df$ABV) == T)
#replace NA ABV values  with mean ABV values
for(i in abv_na_rows){
  final_df$ABV[i] <- final_df$meanABV[i]
}
#find row numbers of na values for ibu and store them
ibu_na_rows <- which(is.na(final_df$IBU) == T)
#replace NA IBU values with mean IBU values
for(i in ibu_na_rows){
  final_df$IBU[i] <- final_df$meanIBU[i]
}
#check to ensure that missing IBU values were not caused by code error
missing_IBU <- final_df %>% filter(is.na(final_df$IBU))
table(missing_IBU$Style)
#remove meanABV and meanIBU columns since they are no longer needed
final_df <- final_df[,-c(11, 12)]
```

##Question 4
Plot the median alcohol content and IBU for each state. Plot a bar chart to compare.

We found the median (middle) value for ABV and IBU for each state. The plots below show how each state compares to the others for both their median ABV and IBU. 

```{r}
#create summary tables for median ABV and IBU values
median_ABV <- final_df %>% group_by(State) %>% summarize(medianABV = median(ABV), count = n()) %>% arrange(State)
median_IBU <- final_df %>% filter(!is.na(final_df$IBU)) %>% group_by(State) %>% summarize(medianIBU = median(IBU), count = n()) %>% arrange(State)
#create plots and plot them one on top of the other
median_ABVPlot <- median_ABV %>% ggplot(mapping = aes(x=State, y=medianABV, fill = State)) + geom_col(show.legend = FALSE) + ylab("Median ABV") + ggtitle("Median ABV by State")
median_IBUPlot <- median_IBU %>% ggplot(mapping = aes(x=State, y=medianIBU, fill = State)) + geom_col(show.legend = FALSE) + ylab("Median IBU") + ggtitle("Median IBU by State")
plot_grid(median_ABVPlot, median_IBUPlot, labels = "AUTO", ncol = 1)
```

##Question 5
Which state has the maximum alcholic (ABV) beer? Which state has the most bitter (IBU) beer?

We found that the Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale from Upslope Brewing Company in Boulder, Colorado had the highest ABV at 12.8%. 
We found that the Bitter Bitch Imperial IPA from Astoria Brewing Company in Astoria, Oregon had the highest IBU at 138. 

```{r}
#return row with max value for ABV and IBU
final_df[which.max(final_df$ABV),]
final_df[which.max(final_df$IBU),]
```
##Question 6
Comment on the summary statistics and distribution of the ABV variable.

The summary shows that the minimum ABV is 0.1%, the maximum is 12.8%, and the median is at 5.65% which is about what we would expect to see. The mean is at 5.98%, so nearly the same as the median. This shows that there are not very many extreme outliers pulling the data in either direction. The histogram and bar plots show similar results with almost 450 beers at 5% ABV, a little less than 400 beers at around 6% ABV, and then a gradual tapering off towards higher ABV values. 
```{r}
#print summary statitics for ABV column
summary(final_df$ABV)
#plot ABV distribution using histogram and boxplot and then plot them side by side. We used Budweiser brand colors for these plots.
abv_hist <- final_df %>% ggplot(mapping = aes(x=ABV)) + geom_histogram(fill="#c8102e") + labs(y = "Count of Beers", title = "Distribution of ABV")
abv_box <- final_df %>% ggplot(mapping=aes(y=ABV)) + geom_boxplot(color="#13294b") + labs(title = "Distribution of ABV") + theme(axis.title.x=element_blank(),axis.text.x=element_blank(), axis.ticks.x=element_blank()) 
plot_grid(abv_hist, abv_box, labels = "AUTO")
```
##Question 7
Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. Make your best judgment of a relationship and EXPLAIN your answer.

It appears that there is a weak relationship between IBU and ABV for beers with IBU values between 0 and 50. Most of the beers here are centered around 5% ABV with a  slight increase in ABV as IBU increases between 0 and 25 and then staying at a fairly consistent ABV for beers between 25 and 50 IBU. For beers with an IBU between 50 and 100 there appears to be more of a linear relationship with ABV increasing more consistently as IBU increases. This relationship tends to flatten out above 100 IBU, but there are fewer beers at this IBU level and this lack of a relationship may simply be a result of a lack of a data. 

```{r}
#draw scatter plot and fit regression line for the relationship between ABV and IBU
final_df %>% ggplot(mapping = aes(x = IBU, y = ABV)) + geom_point(position = "jitter") + geom_smooth(method ="loess") +ggtitle("Relationship Between IBU and ABV")
```

##Question 8
Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually.

We first split the data into trainin and test sets using 70% of the data for training and 30% of the data for testing. We then ran tests using several different values of K and found that a K value of 5 provided the best accuracy. Using IBU and ABV our model could accurately predict whether a beer was an IPA or Ale with approx 91% accuracy. It predicted ales correctly approx 92% of the time and IPAs approx 87% correctly. 

```{r}
#create empty column in the dataframe called Group and fill it with NA
final_df['Group'] <- NA
#find rows with IPA
ipa_rows <- grep("IPA", final_df$Style, ignore.case = T)
#replace NA in group with IPA for IPA rows
for(i in ipa_rows){
  final_df$Group[i] <- "IPA"
}
#find rows with Ale
ale_rows <- grep("Ale", final_df$Style, ignore.case = T)
#replace NA in group with Ale for Ale rows (the is.NA is necessary to avoid over writing IPA styles that also had ale in the name)
for(i in ale_rows){
  if(is.na(final_df$Group[i])==T){
    final_df$Group[i] <- "Ale"
  }
}

#Set the remaining NAs in group as Other
final_df$Group[is.na(final_df$Group)] <- "Other"


#create a datafame that only contains beers in the Ale or IPA group
ABV_data <- final_df %>% filter(Group == "Ale" | Group == "IPA")
ABV_data <- ABV_data %>% filter(!is.na(IBU))

#plot the IBU and ABV data for ABV_Data
ABV_data %>% ggplot(mapping = aes(x = IBU, y = ABV, color = Group)) + geom_point(position = "jitter") + ggtitle("Relationship Between IBU and ABV for Ales and IPAs Only") +
  scale_color_manual(values = c("#c8102e","#13294b"))

#run several tests to find best value of k
splitPerc = .70
iterations = 500
numks = 15
masterAcc = matrix(nrow = iterations, ncol = numks)
for(j in 1:iterations)
{
  accs = data.frame(accuracy = numeric(15), k = numeric(15))
  trainIndices = sample(1:dim(ABV_data)[1],round(splitPerc * dim(ABV_data)[1]))
  train = ABV_data[trainIndices,]
  test = ABV_data[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(3,4)],test[,c(3,4)],train$Group, prob = TRUE, k = i)
    CM = confusionMatrix(table(classifications,test$Group))
    masterAcc[j,i] = CM$overall[1]
  }
  
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
#The best K value was found to be 5. Run 1 knn test with k = 5 and return a sample confusion matrix
classifications = knn(train[,c(3,4)],test[,c(3,4)],train$Group, prob = TRUE, k = 5)
CM = confusionMatrix(table(classifications,test$Group))
CM
#run 100 knn tests with randomly set seed for sampling and knn = 5 to find average accuracy, sensitivity, and specificity 
iterations = 100
masterAcc = matrix(nrow = iterations)
masterSen = matrix(nrow = iterations)
masterSpe = matrix(nrow = iterations)
for(j in 1:iterations)
{
  set.seed(j)
  accs = data.frame(accuracy = numeric(15), k = numeric(15))
  trainIndices = sample(1:dim(ABV_data)[1],round(splitPerc * dim(ABV_data)[1]))
  train = ABV_data[trainIndices,]
  test = ABV_data[-trainIndices,]
  classifications = knn(train[,c(3,4)], test[,c(3,4)], train$Group, prob = TRUE, k = 5)
  CM = confusionMatrix(table(test$Group, classifications))
  masterAcc[j] = CM$overall[1]
  masterSen[j] = CM$byClass[1]
  masterSpe[j] = CM$byClass[2]
}
MeanAcc = colMeans(masterAcc)
masterSen[is.na(masterSen)] = 0 
MeanSen = colMeans(masterSen)
masterSpe[is.na(masterSpe)] = 0
MeanSpe = colMeans(masterSpe)
MeanAcc
MeanSen
MeanSpe
```

##Question 9
Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 

For this question we made tables showing how many of each style of beer is being produced and what the most popular style of beer is in each state. We also plotted the number and style of the top 10 beers being produced and the number of beers being produced for the most popular beer style in each state. 

```{r}
#find how much of each style of beer is being produced in each state
#create dataframe to hold table for count of each style of beer for each state
freq_table <- as.data.frame(table(final_df$State, final_df$Style))
#drop empty rows and 0 occurrence styles (these rows are in error, after this step and summing the frequency column, we do return the correct 2,405 rows)
freq_table <- freq_table %>% filter(Var2 != "")
freq_table <- freq_table %>% filter(Freq != 0)
#arrange by state
freq_table <- freq_table %>% arrange(Var1)
#arrange beers by most being produced to least so we can find the most popular
sum_popbeers <- freq_table %>% group_by(Var2) %>% summarize(sumFreq = sum(Freq))  %>% arrange(desc(sumFreq))
#arrange the data for graphing
sum_popbeers$Var2 <- factor(sum_popbeers$Var2, levels = sum_popbeers$Var2[order(-sum_popbeers$sumFreq)])
#graph how many of each of the top 10 beers are being produced
top10beer_plot <- sum_popbeers[1:10,] %>% ggplot(mapping = aes(x = Var2, y = sumFreq, fill = Var2)) + geom_col(show.legend = FALSE) + geom_text(aes(label = sumFreq),position=position_dodge(width=0.9), vjust=-0.25) +
  scale_fill_brewer(palette="Paired") + labs(title = "Count of Top 10 Styles of Beers Being Produced", x = "Style of Beer", y = "Numbers of Beers of This Style")
#create dataframe where each row reprents most popular beer style for each state
mostpopbeer_state <- freq_table %>% group_by(Var1) %>% slice(which.max(Freq))
#graph most popular beers by state
beer_by_state_plot <- mostpopbeer_state %>% ggplot(mapping = aes(x = Var1, y = Freq, fill = Var2)) + geom_col() + geom_text(aes(label = Freq), position=position_dodge(width=0.9), vjust=-0.25) +
  scale_fill_brewer(palette="Paired") + labs(title = "Count of Most Popular Beers Being Produced By State", x = "State", y = "Numbers of Beers of This Style", fill = "Style of Beer")
#make table column names more human friendly
names(sum_popbeers)[names(sum_popbeers) == "Var2"] <- "Style of Beer"
names(sum_popbeers)[names(sum_popbeers) == "sumFreq"] <- "Number Being Produced"
sum_popbeers
names(mostpopbeer_state)[names(mostpopbeer_state) == "Var1"] <- "State Name"
names(mostpopbeer_state)[names(mostpopbeer_state) == "Var2"] <- "Style of Beer"
names(mostpopbeer_state)[names(mostpopbeer_state) == "Freq"] <- "Number Being Produced"
mostpopbeer_state
top10beer_plot
beer_by_state_plot
```

##Conclusion

In this analysis we reviewed 2,410 beers produced by 558 breweries around the United States. We plotted the number of breweries in each state and created a table showing those numbers as well. We filled in the missing IBU and ABV data with the mean IBU and ABV for each style so that we could perform further analysis. We then plotted the median ABV and IBU for each state.  We found that Colorado produced the beer with  the highest ABV and that Oregon produced the beer with the highest IBU. We showed the distribution of ABV and found that the median ABV is 5.65% with the minimum at 0.1% and the maximum at 12.8%. We found that there may be a weak relationship between ABV and IBU, but mostly for beers with an IBU between 50 and 100. Using kNN we were able to show that knowning the ABV and IBU of a given beer allowed us to predict whether that beer was an IPA or an Ale with roughly 91% accuracy. We then went on to plot the most frequently produced style of beer for each state, so that Budweiser can see what is the most popular beer style in their markets. 

